Convolutional neural networks performing a visual search task show attention-like limits on accuracy when trained to generalize across multiple search stimuli

### 2019-06-04 docs/ccneuro2019/paper.tex
rewrite title and abstract again

\title{Do convolutional neural networks exhibit human-like attentional effects? 
Conflicting results from a visual search task}

@@

\section{Abstract}
{
\bf
Convolutional neural networks (CNNs) trained to perform tasks such as image classification
with human-like accuracy acquire representations similar to those in the primate 
visual system. Do CNNs that are pre-trained to use such representations conversely exhibit 
human-like behavior? We test this with a visual search task that has long been used to 
investigate attention mechanisms. We first show that CNNs using weights trained for image 
classification can exhibit human-like attentional effects when performing this visual 
search task. However, we then demonstrate that these effects can be 
eliminated, without changing the lower-level learned representations, by modifying other 
parameters of training and the balance of classes in the training data. One explanation 
for this finding is that we are only training networks to discriminate single types of 
visual search stimuli. We further test CNNs trained to perform multiple tasks.
We go on to show that manipulating the 
discriminability of the target and distractors in the task can produce attentional 
effects in spite of the changes we made to training. Based on these results, we conclude 
that CNNs do not show attentional effects.
}
\begin{quote}
\small
\textbf{Keywords:} 
attention;visual search
\end{quote}

>>>>

\title{Convolutional neural networks performing a visual search task show attention-like 
limits on accuracy when trained to generalize across multiple search stimuli}

@@

\section{Abstract}
{
\bf
What limits our ability to find what we are looking for in the cluttered noisy world we 
see around us? A now classic form of visual search task has long been used to investigate 
proposed mechanisms of selective attention. In spite of nearly half a century of studies 
using this task, it remains unclear how best to relate different effects found using this 
task to computations that take place in the visual system. A separate thread of research 
has studied the visual system of humans and other primates using convolutional neural 
networks (CNNs) as models. While CNNs have proven quite useful as models of the visual 
system, it also remains unclear the extent to which their behavior mirrors that of 
humans when performing vision-based tasks. Here we investigate whether the accuracy of 
CNNs is affected by the same factors that affect human performance on the classic form of 
visual search task used to investigate attention mechanisms. We first replicate and extend
a previous study which did find that CNNs trained for image classification show human-like
attentional limitations. Then we demonstrate that, by changing how we train the networks, 
these effects can largely be eliminated. Lastly we show that accuracy is impaired when 
single networks are trained to discriminate multiple types of visual search stimuli. Based
on these findings, we suggest that the need to generalize across multiple tasks may be one 
source of limitations on visual search for both CNNs and the primate visual system.
}
\begin{quote}
\small
\textbf{Keywords:} 
attention;visual search
\end{quote}
